Technical Report

### 2018-06-11 fully-abstract-comp/technical-report.tex
Progress on component-related lemmas.

@@

% \dominique{22-1-2018: the below is an alternative proposal to the above definition. It is part of a proposed refactoring to simplify the merge the cases for return pointers and closures in the value relation.}
Two expression relations: one for sealed code-data pairs being jumped to and one for capabilities being jumped to in the regular way.
The argument for having one relation relate pairs of pairs of capabilities and the other relate pairs of capabilities is that that is how xjump and regular jumps work: xjump takes pairs while regular jumps take single capabilities.
\begin{align*}
  \lrexjg{\trust}(W) &= \left\{ \npair{\stpair[.]{v_{c,S},v_{d,S}}{v_{c,T},v_{d,T}}} \middle| 
    \begin{array}{l}
      \forall n' \leq n, \src{\reg_S}, \reg_T, \src{\ms_S}, \ms_T, \src{\ms_\stk}, \src{\stk} \ldotp\\
      \quad \forall W_\lrrs , W_\lrm \ldotp \\
      \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrrg{\trust}(\{\rdata\}) (W_\lrrs ) \wedge\\
      \qquad\memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \wedge \\
      \qquad\Phi_S = \src{(\ms_S,\reg_S,\stk, \ms_\stk)}\wedge\\
      \qquad\Phi_T = (\ms_T,\reg_T) \wedge\\
      \qquad W \oplus W_\lrrs \oplus W_\lrm \text{ is defined }\\
      \qquad\Rightarrow \exists \Phi_S',\Phi_T'\ldotp\\
      \quad\qquad \Phi_S' = \xjumpResult{v_{c,S}}{v_{d,S}}{\Phi_S} \tand\\
      \quad\qquad\Phi_T' = \xjumpResult{v_{c,T}}{v_{d,T}}{\Phi_T}\tand\\
      \quad\qquad\npair[n']{\left(\Phi_S', \Phi_T' \right)}\in \lro
    \end{array}
    \right\}\\
  \lreg{\trust}(W) &= \left\{ \npair{\stpair[.]{v_{c,S}}{v_{c,T}}} \middle| 
    \begin{array}{l}
      \forall n' \leq n, \src{\reg_S}, \reg_T, \src{\ms_S}, \ms_T, \src{\ms_\stk}, \src{\stk} \ldotp\\
      \quad \forall W_\lrrs , W_\lrm \ldotp \\
      \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrrg{\trust} (W_\lrrs ) \wedge\\
      \qquad\memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \\
      \qquad\Phi_S = \src{(\ms_S,\reg_S,\stk, \ms_\stk)}\\
      \qquad \Phi_S' = \Phi_S \updReg{\pcreg}{v_{c,S}}\\
      \qquad\Phi_T = (\ms_T,\reg_T)\\
      \qquad\Phi_T' = \Phi_T\updReg{\pcreg}{v_{c,T}}\\
      \qquad W \oplus W_\lrrs \oplus W_\lrm\\
      \qquad\Rightarrow\npair[n']{\left(\Phi_S', \Phi_T' \right)}\in \lro
    \end{array}
    \right\}
\end{align*}

@@

\[
  \lrec(W) = \left\{ \npair{\left(
        \arraycolsep=0pt\array{l}
        (\ms_S,\reg_S,\stk,\ms_\stk),\\
        (\ms_T,\reg_T)\endarray\right)} \middle|
    \begin{array}{l}
      \gc = (\ta,\stkb) \tand\\
      \exists W_M,W_R,W_\pcreg \ldotp W = W_M \oplus W_R \oplus W_\pcreg \tand\\
      \reg_S(\pcreg) = ((\rx,\lin),\baddr,\eaddr,\aaddr) \tand (\trust = \trusted \text{ iff } \aaddr \in \ta) \tand \\
      \npair{( \reg_S(\pcreg),\reg_T(\pcreg) )} \in \lrvg{\trust}(W_\pcreg) \tand \\
      \memSat{\ms_S,\ms_\stk,\stk,\ms_T}{W_M} \tand\\
      \npair{\stpair{\reg}{\reg}} \in \lrrg{\trust}(W_R)
    \end{array}
            \right\}
\]

\begin{lemma}[Compatibility lemma for initial execution configuration construction]
  If
  \begin{itemize}
  \item $\npair{\stpair{\var{comp}}{\var{comp}}} \in \lrcomp(W)$
  \item $\var{comp} \rightsquigarrow \Phi$,
  \item TODO: no stack regions yet?
  \end{itemize}
  then $\exists W' \future W$ such that
  \begin{itemize}
  \item $\npair{\stpair{\Phi}{\Phi}}\in\lrec(W')$
  \end{itemize}
\end{lemma}
\begin{proof}
  Take $\var{comp} = ((\mscode, \msdata, \overline{c_{\mathrm{import}}}, \overline{c_{\mathrm{export}}}, \sigrets, \sigcloss,A_\linear),  c_{\mathrm{main},c}, c_{\mathrm{main},d})$.
  Then we know from $\var{comp} \rightsquigarrow \Phi$ that
  \begin{itemize}
  \item $\reg(\pcreg) = c_{\mathrm{main},c}$
  \item $\reg(\rdata) = c_{\mathrm{main},d}$ 
  \item $\src{\reg(\rstk) = \stkptr{\rx,\baddr_\stk,\eaddr_\stk,\eaddr_\stk}}$ 
  \item $\trg{\reg(\rstk) = ((\rx,\linear),\baddr_\stk,\eaddr_\stk,\eaddr_\stk)}$ 
  \item $\reg(\RegName \setminus \{\pcreg,\rdata,\rstk\}) = 0$
  \item $\range{\ms_\stk} = \{0\}$
  \item $\mem = \mscode \uplus \msdata \trg{\;\uplus\; \ms_\stk}$
  \item $[\baddr_\stk,\eaddr_\stk] = \dom(\ms_\stk ) \mathrel{\#} (\dom(\mscode) \cup \dom(\msdata))$
  \item $\overline{c_{\mathrm{import}}} = \emptyset$
  \end{itemize}

  From $\npair{\stpair{\var{comp}}{\var{comp}}} \in \lrcomp(W)$, we know that
  $\{c_{\mathrm{main},c}, c_{\mathrm{main},d}\}\subseteq \overline{c_{\mathrm{export}}}$ and 
  $\npair{(\var{comp}_0,\var{comp}_0)} \in \lrcomp(W)$ with $\var{comp}_0 = (\mscode, \msdata, \overline{c_{\mathrm{import}}}, \overline{c_{\mathrm{export}}}, \sigrets, \sigcloss,A_\linear)$.
  From this, it follows that for all  $W' \future W$, if
  \begin{equation*}
    \overline{\npair[n']{(c_{\mathrm{import}},c_{\mathrm{import}})}} \in \lrv(\purePart{W'}) \text{ for all $n' < n$}
  \end{equation*}
  then
  \begin{equation*}
    \overline{\npair{(c_{\mathrm{export}},c_{\mathrm{export}})}} \in \lrv(\purePart{W'})
  \end{equation*}
  Since $\overline{c_{\mathrm{import}}} = \emptyset$, the former holds vacuously and the latter follows for every $W' \future W$.

  Now take $\overline{r_a} \not\in\dom(\pwfree)$ and 
  \begin{align*}
    W' &= W\overline{\update{\mathrm{free}(r_a)}{\stdreg{\{a\},\gc}{\pur}}}_{a \in [\baddr_\stk,\eaddr_\stk]}\\
   W_\pcreg &= \purePart{W'}\\
    W_R &= \purePart{W}\overline{\update{\mathrm{free}(r_a)}{\stdreg{\{a\},\gc}{\spatialo}}}_{a \in [\baddr_\stk,\eaddr_\stk]}\\
    W_M &= W\overline{\update{\mathrm{free}(r_a)}{\stdreg{\{a\},\gc}{\spatial}}}_{a \in [\baddr_\stk,\eaddr_\stk]}
  \end{align*}
  Then $W' = W_M \oplus W_R \oplus W_\pcreg$ and by definition of $\lrec$, it suffices to prove that 
  \begin{itemize}
  \item $\reg_S(\pcreg) = ((\rx,\lin),\baddr,\eaddr,\aaddr)$:
    TODO
  \item $(\trust = \trusted \text{ iff } \aaddr \in \ta)$:
    TODO
  \item $\npair{( \reg_S(\pcreg),\reg_T(\pcreg) )} \in \lrvg{\trust}(W_\pcreg)$:
    TODO
  \item $\memSat{\ms_S,\ms_\stk,\stk,\ms_T}{W_M}$:
    TODO
  \item $\npair{\stpair{\reg}{\reg}} \in \lrrg{\trust}(W_R)$:
    TODO
  \end{itemize}

\end{proof}

\begin{lemma}[Adequacy of logical relation]
  If
  \begin{itemize}
  \item $\npair{\stpair{\Phi}{\Phi}}\in\lrec[\preceq](W)$
  \item $i \leq n$
  \item $\Phi_S \sterm[i]{\gc}$
  \end{itemize}
  then $\Phi_T\term$.

>>>>

\begin{lemma}[purePart respects $\oplus$]
  \label{lem:purePart-oplus}
 For all $W_1,W_2$, we have that $\purePart{W_1\oplus W_2} = \purePart{W_1} \oplus \purePart{W_2}$.
\end{lemma}
\begin{proof}
  Follows from the definition of $\purePart{}$ and $\oplus$.
\end{proof}

@@

% \dominique{22-1-2018: the below is an alternative proposal to the above definition. It is part of a proposed refactoring to simplify the merge the cases for return pointers and closures in the value relation.}
Two expression relations: one for sealed code-data pairs being jumped to and one for capabilities being jumped to in the regular way.
The argument for having one relation relate pairs of pairs of capabilities and the other relate pairs of capabilities is that that is how xjump and regular jumps work: xjump takes pairs while regular jumps take single capabilities.
\begin{align*}
  \lrexj(W) &= \left\{ \npair{\stpair[.]{v_{c,S},v_{d,S}}{v_{c,T},v_{d,T}}} \middle| 
    \begin{array}{l}
      \forall n' \leq n, \src{\reg_S}, \reg_T, \src{\ms_S}, \ms_T, \src{\ms_\stk}, \src{\stk} \ldotp\\
      \quad \forall W_\lrrs , W_\lrm \ldotp \\
      \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrrg{\untrusted}(\{\rdata\}) (W_\lrrs ) \wedge\\
      \qquad\memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \wedge \\
      \qquad\Phi_S = \src{(\ms_S,\reg_S,\stk, \ms_\stk)}\wedge\\
      \qquad\Phi_T = (\ms_T,\reg_T) \wedge\\
      \qquad W \oplus W_\lrrs \oplus W_\lrm \text{ is defined }\\
      \qquad\Rightarrow \exists \Phi_S',\Phi_T'\ldotp\\
      \quad\qquad \Phi_S' = \xjumpResult{v_{c,S}}{v_{d,S}}{\Phi_S} \tand\\
      \quad\qquad\Phi_T' = \xjumpResult{v_{c,T}}{v_{d,T}}{\Phi_T}\tand\\
      \quad\qquad\npair[n']{\left(\Phi_S', \Phi_T' \right)}\in \lro
    \end{array}
    \right\}\\
  \lreg{\trust}(W) &= \left\{ \npair{\stpair[.]{v_{c,S}}{v_{c,T}}} \middle| 
    \begin{array}{l}
      \forall n' \leq n, \src{\reg_S}, \reg_T, \src{\ms_S}, \ms_T, \src{\ms_\stk}, \src{\stk} \ldotp\\
      \quad \forall W_\lrrs , W_\lrm \ldotp \\
      \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrrg{\trust} (W_\lrrs ) \wedge\\
      \qquad\memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \\
      \qquad\Phi_S = \src{(\ms_S,\reg_S,\stk, \ms_\stk)}\\
      \qquad \Phi_S' = \Phi_S \updReg{\pcreg}{v_{c,S}}\\
      \qquad\Phi_T = (\ms_T,\reg_T)\\
      \qquad\Phi_T' = \Phi_T\updReg{\pcreg}{v_{c,T}}\\
      \qquad W \oplus W_\lrrs \oplus W_\lrm\\
      \qquad\Rightarrow\npair[n']{\left(\Phi_S', \Phi_T' \right)}\in \lro
    \end{array}
    \right\}
\end{align*}

@@

\[
  \lrec(W) = \left\{ \npair{\left(
        \arraycolsep=0pt\array{l}
        (\ms_S,\reg_S,\stk,\ms_\stk),\\
        (\ms_T,\reg_T)\endarray\right)} \middle|
    \begin{array}{l}
      \gc = (\ta,\stkb) \tand\\
      \exists W_M,W_R,W_\pcreg \ldotp W = W_M \oplus W_R \oplus W_\pcreg \tand\\
      \npair{( (\reg_S(\pcreg),\reg_S(\rdata)), (\reg_T(\pcreg),\reg_T(\rdata)) )} \in \lrexj(W_\pcreg) \tand \\
      \memSat{\ms_S,\ms_\stk,\stk,\ms_T}{W_M} \tand\\
      \npair{\stpair{\reg}{\reg}} \in \lrrg{\trust}(\{\rdata\})(W_R)
    \end{array}
            \right\}
\]

\begin{lemma}[Compatibility lemma for initial execution configuration construction]
  If
  \begin{itemize}
  \item $\npair{(\var{comp},\var{comp})} \in \lrcomp(W)$
  \item $\dom(\pwfree) = \dom(\pwpriv) = \emptyset$
  \item $\var{comp} \rightsquigarrow \Phi$,
  \end{itemize}
  then $\exists W' \future W$ such that
  \begin{itemize}
  \item $\npair{(\Phi,\Phi)}\in\lrec(W')$
  \end{itemize}
\end{lemma}
\begin{proof}
  Take $\var{comp} = ((\mscode, \msdata, \overline{c_{\mathrm{import}}}, \overline{c_{\mathrm{export}}}, \sigrets, \sigcloss,A_\linear),  c_{\mathrm{main},c}, c_{\mathrm{main},d})$.
  Then we know from $\var{comp} \rightsquigarrow \Phi$ that
  \begin{itemize}
  \item $\reg(\pcreg) = c_{\mathrm{main},c}$
  \item $\reg(\rdata) = c_{\mathrm{main},d}$ 
  \item $\src{\reg(\rstk) = \stkptr{\rx,\baddr_\stk,\eaddr_\stk,\eaddr_\stk}}$ 
  \item $\trg{\reg(\rstk) = ((\rx,\linear),\baddr_\stk,\eaddr_\stk,\eaddr_\stk)}$ 
  \item $\reg(\RegName \setminus \{\pcreg,\rdata,\rstk\}) = 0$
  \item $\range{\ms_\stk} = \{0\}$
  \item $\mem = \mscode \uplus \msdata \trg{\;\uplus\; \ms_\stk}$
  \item $[\baddr_\stk,\eaddr_\stk] = \dom(\ms_\stk ) \mathrel{\#} (\dom(\mscode) \cup \dom(\msdata))$
  \item $\overline{c_{\mathrm{import}}} = \emptyset$
  \item $\Phi = (\mem, \reg\src{, \emptyset, \ms_\stk})$
  \end{itemize}

  From $\npair{(\var{comp},\var{comp})} \in \lrcomp(W)$, we know that
  $\{c_{\mathrm{main},c}, c_{\mathrm{main},d}\}\subseteq \overline{c_{\mathrm{export}}}$ and 
  $\npair{(\var{comp}_0,\var{comp}_0)} \in \lrcomp(W)$ with $\var{comp}_0 = (\mscode, \msdata, \overline{c_{\mathrm{import}}}, \overline{c_{\mathrm{export}}}, \sigrets, \sigcloss,A_\linear)$.
  From this, it follows that for all  $W' \future W$, if
  \begin{equation*}
    \overline{\npair[n']{(c_{\mathrm{import}},c_{\mathrm{import}})}} \in \lrv(\purePart{W'}) \text{ for all $n' < n$}
  \end{equation*}
  then
  \begin{equation*}
    \npair{(\sigrets\uplus\sigcloss,\mscode\uplus \msdata, \mscode\uplus\msdata)} \in \lrheap(\pwheap)(W')
  \end{equation*}
  and
  \begin{equation*}
    \overline{\npair{(c_{\mathrm{export}},c_{\mathrm{export}})}} \in \lrv(\purePart{W'})
  \end{equation*}
  Since $\overline{c_{\mathrm{import}}} = \emptyset$, the former holds vacuously and the latter two results follow for every $W' \future W$.

  Now take $\overline{r_a}$ for $a \in [\baddr_\stk,\eaddr_\stk]$ arbitrary and 
  \begin{align*}
    \pwheap[W_\stk] &= \emptyset\\
    \pwpriv[W_\stk] &= \emptyset\\
    \dom(\pwfree[W_\stk]) &= \{r_a \mid a \in [\baddr_\stk,\eaddr_\stk]\}\\
    \pwfree[W_\stk](r_a) &= \stdreg{\{a\},\gc}{\spatialo}\\
    W' &= W \uplus W_\stk\\
    W_\pcreg &= \purePart{W'}\\
    W_R &= \purePart{W} \uplus W_\stk\\
    W_M &= W \uplus \purePart{W_\stk}
  \end{align*}
  Take $\trust = \trusted$ iff $a \in \ta$. 
  Then $W' = W_M \oplus W_R \oplus W_\pcreg$ (by Lemmas~\ref{lem:purePart-duplicable},~\ref{lem:oplus-distr-uplus},~\ref{lem:oplus-assoc-comm} and~\ref{lem:uplus-assoc-comm}), $W_M \future W$ (by Lemma~\ref{lem:uplus-future}) and $W_\pcreg \future \purePart{W}$ (by Lemma~\ref{lem:uplus-future} and~\ref{lem:purePart-oplus}).

  By definition of $\lrec$, it suffices to prove that 
  \begin{itemize}
  \item 
    $\npair{( (\reg_S(\pcreg),\reg_S(\rdata)), (\reg_T(\pcreg),\reg_T(\rdata)) )} \in \lrexj(W_\pcreg)$:
    TODO
  \item $\memSat{\ms_S,\ms_\stk,\stk,\ms_T}{W_M}$:
    Since $W_M \future W$, we have seen above that 
    \begin{equation*}
      \npair{(\sigrets\uplus\sigcloss,\mscode\uplus \msdata, \mscode\uplus\msdata)} \in \lrheap(\pwheap[W_M])(W_M)
    \end{equation*}
    It suffices to prove that also
    \begin{itemize}
    \item $\memSatStack{\emptyset,\emptyset}{\purePart{W_M}}}$
    \item $\memSatFStack{\ms_\stk,\ms_\stk}{\purePart{W_M}}$
    \end{itemize}
    The former follows vacuously.
    The latter follows by taking $R_\ms(r_a) = (\ms_\stk|_{\{a\}},\ms_\stk|_{\{a\}})$, $R_W(r_a) = \purePart{W_M}$ for $a \in [\baddr_\stk,\eaddr_\stk]$ by Lemma~\ref{lem:purePart-duplicable} and~\ref{lem:purePart-idempotent} if we can show that:
    \begin{equation*}
      \npair{(\ms_\stk|_{\{a\}},\ms_\stk|_{\{a\}})} \in \stdreg{\{a\},\gc}{\spatialo}.H~ \xi^{-1}(\purePart{W_M})
    \end{equation*}
    By definition, it suffices to show that for all $n' < n$, we have that:
    \begin{equation*}
      \npair{(\ms_\stk(a),\ms_\stk(a)) \in \lrv(\purePart{W_M})}
    \end{equation*}
    But $\ms_\stk(a) = 0$, so this follows easily by definition.
  \item $\npair{\stpair{\reg}{\reg}} \in \lrrg{\trust}(W_R)$:
    TODO
  \end{itemize}
\end{proof}

\begin{lemma}[Adequacy of execution configuration LR]
  If
  \begin{itemize}
  \item $\npair{\stpair{\Phi}{\Phi}}\in\lrec[\preceq](W)$
  \item $i \leq n$
  \item $\Phi_S \sterm[i]{\gc}$
  \end{itemize}
  then $\Phi_T\term$.
